Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression\thanksref{T1}

### 2019-06-21 revision_2/paper.tex
created new plots

\response{To asses the predictive capability, we seek a metric that takes into account both the estimation of $\theta_{i}$ and $\sigma^{2}$. In this model, estimation of $\theta_{i}$ and prediction are analogous. To this end we consider the expected value of  the log likelihood ratio between the true distribution of good data and a model estimate of this distribution
\begin{equation}
\label{sim1loss}
E[\log f(y|\theta_{i}, \sigma^{2}) - \log f(y|\hat{\theta_{i}}, \hat{\sigma^{2}})] =
\log(\hat\sigma^{2}) - \log(\sigma^{2}) + \frac{1}{2\hat\sigma^{2}}(\sigma^{2} + (\theta - \hat{\theta_{i}})^{2})
\end{equation}
The expectation is taken with respect to the true data distribution of the good data assuming fixed estimates. These estimates are the posterior means for the Bayesian methods and the standard point estimates for the classical methods.
}

@@

\response{STILL NEED TO COMPLETE}
In this simulation the data are generated from the following mechanism: $y = \bbeta^{\top}x + \epsilon$
where $\bbeta = (\beta_{1}, \beta_{2},\beta_{3})^{\top}$ with $\beta_{i} \iid N(0,1)$ and  $x \sim Uniform(0,1)$. The error $\epsilon \sim N(0,\sigma^{2})$ with probability $0.9$ and $\epsilon \sim Half-Normal(0,10\sigma^{2})$ with probability $0.1$. Finally, the variance $\sigma^{2} \sim IG(5,5)$. Thus, there are three active covariates, with one-sided outlier contamination. The regression model used for fitting contains 17 additional covariates generated uniformly on (0,1) representing the situation where several potential covariates are available and a goal is variable selection. For the simulation, $K = 50$ data sets of size $n = 500$ are generated from the true model and models including the additional non-active covariates are fit. We consider the classical robust regression using Tukey's estimator of location and Huber's estimator of scale, along with a corresponding restricted likelihood version. For the priors, denote $\bbeta_{all} = (\bbeta, \bbeta_{extra})^{\top}$ be the active coefficients augmented with the non-active extra coefficients. We take $\bbeta_{all}\sim N_{20}(\mb{0}, \sigma^{2}_{\beta}I)$ and $\sigma^{2}\sim IG(5,5)$. Hence, the prior for $\sigma^{2}$ is correct and the prior for $\bbeta$ is at least centered at the correct value as well. We also consider the Student-t model with compariable priors and fixed degrees of freedom $\nu = 5$.  For each data set, we fit the models for $ \sigma_{\beta} = 0.2, 0.4, 0.6, 0.8,,1.0$. The acceptance rates for the restricted likelihood MCMC data-augmentation step range from ADDDDDDDD.

To compare estimation performance we first consider the squared error SE$_{k} = ||\bbeta - \hat\bbeta||^{2} + ||\hat\bbeta_{extra}||^{2}$ for each simulation $k$. The mean squared error and standard errors over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{mseSimMany}.

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{mse_sim_many_p.png}
\caption{ADD}
\label{mseSimMany}
\end{figure}

Additionally, we consider the log marginal distributions to understand predictive performance....
FINISH

>>>>

\response{To asses the predictive capability, we seek a metric that takes into account both the estimation of $\theta_{i}$ and $\sigma^{2}$. In this model, estimation of $\theta_{i}$ and prediction are analogous. To this end we consider
\begin{equation}
\label{sim1loss}
E[\log f(y|\theta_{i}, \sigma^{2}) - \log f(y|\hat{\theta_{i}}, \hat{\sigma}^{2})] =
\log(\hat\sigma^{2}) - \log(\sigma^{2}) + \frac{1}{2\hat\sigma^{2}}(\sigma^{2} + (\theta - \hat{\theta_{i}})^{2})
\end{equation}
The expectation is taken with respect to the true data distribution of the good data assuming fixed estimates. These estimates are the posterior means for the Bayesian methods and the standard point estimates for the classical methods.
}

@@

In this simulation the data are generated from the following mechanism: $y = \bbeta^{\top}x + \epsilon$
with $\bbeta = (\beta_{1}, \beta_{2},\beta_{3})^{\top}$ and the error $\epsilon \sim N(0,\sigma^{2})$ with probability $0.8$ and $\epsilon \sim \text{Half-Normal}(0,25\sigma^{2})$ with probability $0.2$ (i.e., there is a relatively large amount of one-sided outlier contamination). The components of $x = (x_{1}, x_{2}, x_{3})$ are correlated with $x_1 \sim N(0,1)$ and $x_{j} = x_{1} + \eta_{j}$ with $\eta_{j} \sim N(0, 4)$ for $j = 2,3$. This results in a theoretical correlation of $1/sqrt(5) \approx 0.44$ between $x_{1}$ and both $x_{j}, j = 2, 3$. The model used for fitting contains an additional 27 covariates, some of which are also correlated with $x_{1}$, $x_{2}$, and $x_{3}$. Specifically the fitting model is $y = \beta^{\top} x + \beta^{*\top} x^{*} + \epsilon$ where $x^{*}$ and $\beta^{*}$ are 27 dimensional vectors of extra covariates and slope parameters. Of these 27 covariates, 21 are generated independently from standard normal distributions. Of the remaining 6, two each are generated by adding standard normal noise to $x_{1}$, $x_{2}$, and $x_{3}$. This represents a common situation where several covariates with various levels of correlation amongst them are available for fitting, but only a few govern the data generating mechanism.

For the simulation, $K = 30$ data sets (including the additional covariates) of size $n = 500$ are generated from the true model with true values $\bbeta = (1,1,1)^{\top}$ and $\sigma^{2} = 2$. We fit the model including all 30 covariates and consider the following methods for the fit 1) classical robust regression with Tukey's estimator of location and Huber's estimator of scale, 2) the corresponding restricted likelihood version 3) A heavy-tailed Bayesian model with a Student-t likelihood with $\nu = 5$ degrees of freedom.  For the Bayesian models we take $\bbeta_{all}\sim N_{20}(\mb{0}, \sigma^{2}_{\beta}I)$ with $\bbeta_{all} = (\bbeta, \bbeta^{*})^{\top}$ and $\sigma^{2}\sim IG(5,8)$ under the restricted model and  $\sigma^{2}\sim IG(5,\frac{\nu-2}{\nu}8)$ under the Student-t model.  For each data set, we fit the models for $ \sigma_{\beta} = 0.4, 0.6, 0.8,\dots, 1.4$. The acceptance rates for the restricted likelihood MCMC data-augmentation step range from $0.3$ to $0.36$ across all the data sets and values of $\sigma_{\beta}$. To compare performance we first consider the $MSE = (||\bbeta - \hat\bbeta||^{2} + ||\hat\bbeta^{*}||^{2})/30$ for each simulation where $\hat\bbeta$ and $\hat\bbeta^{*}$ are point estimates for the fitted model. For the Bayesian models, we use posterior means. The average MSE plus/minus one standard error over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{mseSimMany}. The classical fit is labeled `rlm` and is the same for each value of the prior standard deviation $\sigma^{2}_{\beta}$. We see for most values of the the prior standard deviation, the Bayesian models (`restricted' and `t') outperform the classical fit. The correlation amongst the covariates causes a certain level of confounding and the prior shrinkage helps to improve estimation. However, too much shrinkage can be detrimental as demonstrated for $\sigma_{\beta} = 0.4$. While this will help for estimation of $\bbeta^{*} = 0$, the estimation of the active parameters $\bbeta$ can be hindered. The $t$ model seems more sensitive to this effect than the restricted model. The restricted model also has an additional advantage when it comes to prediction of the non-outlying data. To see this, for each simulation we consider the mean negative log-likelihood of the non-outlying data: $MNLL = -\frac{1}{N} \sum \log f(y_{i} |  \hat\bbeta, \hat\bbeta^{*}, \hat\sigma)$ where $f$ is the assumed likelihood and the average is taken over the $N$ non-outlying points $y_{i}$. For the classical and restricted fits, $f$ is the normal likelihood and for the `t' it is the heavy-tailed Student-t likelihood. The average MNLL plus/minus one standard error over the simulations for each $\sigma^{2}_{\beta}$ are displayed in Figure \ref{negllSimMany}. First, the restricted version has a small but consistent improvement over the classical method. Second, it is clear that that the heavy-tailed model suffers when trying to predict the non-outlying data since it assumes the entire data generating mechanism is heavy-tailed.

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{mse_sim_many_p.png}
\caption{Average MSE plus/minus one standard error over the $K = 30$ simulations for each value of the prior standard deviation ($\sigma^{2}_{\beta}$) and each of the fitting methods. `Restricted' is our method conditioning on Tukey's estimator of location and Huber's estimator of scale. `rlm' refers to the classical robust linear model fit with the same estimators and `t' is the heavy-tailed Bayesian model with a Student-t likelihood. The `rlm' results are the same for each $\sigma^{2}_{\beta}$.}
\label{mseSimMany}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width = 5in]{negll_sim_many_p.png}
\caption{Average MNLL plus/minus one standard error over the $K = 30$ simulations for each value of the prior standard deviation ($\sigma^{2}_{\beta}$) and each of the fitting methods. `Restricted' is our method conditioning on Tukey's estimator of location and Huber's estimator of scale. `rlm' refers to the classical robust linear model fit with the same estimators and `t' is the heavy-tailed Bayesian model with a Student-t likelihood. The `rlm' results are the same for each $\sigma^{2}_{\beta}$.}
\label{negllSimMany}
\end{figure}
